---
title: "Consolidating regression variables for EU identity inference"
author: "Balázs Dukai (<balazs.dukai@gmail.com>) [www.linkedin.com/in/balazsdukai](www.linkedin.com/in/balazsdukai)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
    number_sections: yes
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes
  html_document: default
subtitle: COHESIFY - The Impact of Cohesion Policy on EU Identification and is part
  of Horizon 2020 program
papersize: a4paper
---

```{r setup, include=FALSE, message=FALSE}
library(haven)
library(xlsx)
library(pander)
library(RPostgreSQL)
library(tidyverse)
library(sf)
library(ggplot2)
db <- dbConnect(RPostgreSQL::PostgreSQL(), dbname = "nuts", user="bdukai", host="localhost")
knitr::opts_chunk$set(connection = "db")
options(knitr.table.format ='pandoc')
```

\pagebreak


```{r test-main, child = '00_nuts.Rmd'}
```

\pagebreak

```{r test-main, child = '01_border_typology.Rmd'}
```

\pagebreak

```{r test-main, child = '02_trust.Rmd'}
```

\pagebreak

```{r test-main, child = '03_qog.Rmd'}
```

\pagebreak

```{r test-main, child = '04_struct.Rmd'}
```

\pagebreak

```{r test-main, child = '05_unemp.Rmd'}
```

\pagebreak

```{r test-main, child = '06_education.Rmd'}
```

\pagebreak

```{r test-main, child = '07_soc_progr.Rmd'}
```

\pagebreak

```{r test-main, child = '08_migration.Rmd'}
```

\pagebreak

```{r test-main, child = '09_access_to_soc.Rmd'}
```

\pagebreak

```{r test-main, child = '10_agriculture.Rmd'}
```

\pagebreak

```{r test-main, child = '11_eu_id.Rmd'}
```

\pagebreak

```{r test-main, child = '12_eu_attachment.Rmd'}
```

\pagebreak

```{r join_everything2, include=FALSE}
eurobarometer_2015 <- nt_eu_2015 %>% 
    left_join(gdp_nuts2, by = "id2.2013") %>% 
    left_join(acc2soc_clean, by = "id2.2013") %>% 
    left_join(struct_2014_clean, by = "id2.2013")

eurobarometer_2016 <- nt_eu_2016 %>% 
    left_join(gdp_nuts2, by = "id2.2013") %>% 
    left_join(acc2soc_clean, by = "id2.2013") %>% 
    left_join(struct_2014_clean, by = "id2.2013")
```

```{r}
write_csv(eurobarometer_2015,
          path = "~/Development/cohesify/cohesify_variables/eu_2015_ZA6643.csv")
write_csv(eurobarometer_2016,
          path = "~/Development/cohesify/cohesify_variables/eu_2016_ZA6694.csv")
```

```{r join_everything, include=FALSE, eval=FALSE}
eu_vars <- typ_border_clean %>% 
    full_join(trust_clean, by = "id2.2013") %>% 
    full_join(qog_clean, by = "id2.2013") %>% 
    full_join(struct_2007_clean, by = "id2.2013") %>% 
    full_join(struct_2014_clean, by = "id2.2013") %>%
    full_join(unemp_clean, by = "id2.2013") %>%
    full_join(edu_clean, by = "id2.2013") %>%
    full_join(soc_clean, by = "id2.2013") %>%
    full_join(migr_clean, by = "id2.2013") %>%
    full_join(acc2soc_clean, by = "id2.2013") %>%
    full_join(agr_clean, by = "id2.2013") %>%
    full_join(eu_id_clean, by = "id2.2013") %>%
    full_join(eu_att_clean, by = "id2.2013")
# %>% 
#     full_join(select(nuts_2013_weight, id2.2013, weight), by = "id2.2013")
```

```{r select_vars, include=FALSE}
# # Variable weigthing
# eu_vars <- eu_vars %>% 
#     mutate_at(c("eqi_2013", "sgi_aggregated"),
#               funs(. * weight))

# Remove regions with missing values for the dependent variable (eu_id and eu_att);
# all regions included, even those that contain missing values;
eu_vars_ri <- eu_vars %>%
    filter(eu_id.incl == 1 & eu_att.incl == 1)
# %>% 
#     select(-ends_with(".incl"))

# these regions are removed
eu_vars_ri_n <- eu_vars %>% 
    filter(eu_id.incl == 0 | eu_att.incl == 0) %>% 
    select(id2.2013)
write_csv(eu_vars_ri, path = "/home/bdukai/Development/cohesify/cohesify_variables/eu_all.csv")

# remove regions that are missing value for any variable;
eu_vars_no_na <- eu_vars_ri[complete.cases(eu_vars_ri), ]
# these regions are removed;
eu_vars_no_na_n <- eu_vars_ri[!complete.cases(eu_vars_ri), "id2.2013"]
write_csv(eu_vars_no_na, path = "/home/bdukai/Development/cohesify/cohesify_variables/eu_no_missing.csv")

# remove all regions where the value for the NUTS2 region is a generalization of
# NUTS0/1 values, because of missing data on the NUTS2 level;
eu_vars_include_n <- eu_vars %>% 
    filter(typ_border.incl == 1 &
           trust.incl == 1 &
           eqi.incl == 1 &
           struct_2007.incl == 1 &
           struct_2014.incl == 1 &
           unemp.incl == 1 &
           edu.incl == 1 &
           soc_idx.incl == 1 &
           migration.incl == 1 &
           sgi_aggregated.incl == 1 &
           agr.incl == 1 &
           eu_id.incl == 1 &
           eu_att.incl == 1) %>% 
    select(-ends_with(".incl"))
# these regions are removed;
eu_vars_exclude <- setdiff(nuts_2013_levels$id2, eu_vars_include_n$id2.2013)
write_csv(eu_vars_include_n, path = "/home/bdukai/Development/cohesify/cohesify_variables/eu_include.csv")
```

# Data quality

Besides the issues specific to each variable, these issues are generally present at most/all variables:

+ The NUTS 2013 level 2 region *FRA5* is a new region, thus it has no corresponding region in the earlier NUTS versions. Therefore the variables have missing values in case they are originally mapped to NUTS regions of 2006/2010 version.

+ Between NUTS versions 2010 and 2013 the NUTS3 region *SI035* (in *SI03*) is a merge of *SI021* (in *SI02*) and *SI015* (in *SI01*). Eight out of nine of the 2010 level 3 regions corresponding to the 2013 level 2 region *SI03* were part of the *SI01* 2010 level 2 region. The only exception is *SI021* (from *SI02*). At the time of the study there was no population data available for NUTS 2010 regions that could be used to weight the variables that are translated from NUTS2/3 2010 to NUTS2 2013. Therefore the 2013 region *SI03* is considered to be a name change from *SI01* to *SI03*, ignoring the influence of *SI021*. This affects the variables `sgi_aggregated, migration_2013, eu_id_type_code, soc_idx`.

# Output

Due to the large number of NUTS2 regions with missing values at many variables, three outputs are prepared:

**`eu_all.csv`**

All regions included, even those that contain missing values for the independent variables. Regions are excluded that are missing values for either of the dependent variables `EU identificaion` or `EU attachment`.

Contains 261/276 NUTS2 regions. Excluded regions are:

```{r, echo=FALSE}
eu_vars_ri_n$id2.2013
```

**`eu_no_missing.csv`**

Regions are removed that are missing value for any variable.

Contains 251/276 NUTS2 regions. Excluded regions are those listed in the previous output and:

```{r, echo=FALSE}
eu_vars_no_na_n
```

**`eu_include.csv`**

Removed all regions where the value for the NUTS2 region is a generalization of NUTS0/1 values, because of missing data on the NUTS2 level.

Contains 125/276 NUTS2 regions. Excluded regions are:

```{r, echo=FALSE}
eu_vars_exclude
```


