# Territorial typology à border regions

**Data source:** Programme documents of *Directorate-General for Regional Policy, Enlargement and External Relations*; For the full description of the study see the Regional Focus paper [n° 01/2011 : Regional typologies: a compilation](http://ec.europa.eu/regional_policy/archive/information/focus/index_en.cfm)

**Date of survey:** 2007-2013

**Nomenclature:** NUTS 2006 level 3

**Variables used:** `border regions`

Variable encoding in the original data set:

+ 0 – No program area
+ 1 – Programme area
+ 9 – Programme area (partly)

As described in Regional Focus No. 1/2011. The NUTS 3-based selection of border regions refers to the regions participating in the core areas of cross-border cooperation programmes in the programming period 2007-2013. Due to NUTS boundary changes, some current NUTS 3 regions are only partly eligible as programme areas.

This variable will become *regions that **participated** in the core areas of cross-border cooperation programmes in the programming period 2007-2013*. Because the values are from 2009, thus they use NUTS 2006 (not 2010), and they refer to the programming period 2007-2013.

## Data processing

1. Join the NUTS3 2006 variable to the NUTS2-3 2010 regions using the 2006 <-> 2010 translation schema, then translate to NUTS 2013.

2. The aggregation of NUTS3 to NUTS2 is achieved by computing the precentage of *No programme / Programme / Programme (partly)* NUTS3 regions in the corresponding NUTS2 2013 regions.

 
```{r border_in, include=FALSE, cache=TRUE}
# Read data
typo_border_regions <- read.xlsx("~/Data/COHESIFY/Datasets for regression/Territorial typology/Typologies_metadata_data_final (2).xls",
                                 sheetName = "data regional typologies",
                                 colIndex = c(1,8), endRow = 1464, header = TRUE,
                                 stringsAsFactors = FALSE) %>% 
    filter(typ_border_A != 2009)
```

```{r join_2006_2010, include=FALSE}
# Join NUTS2-3 2010, NUTS3 2006, typ_border_A
typo_border_regions <- typo_border_regions %>% 
    full_join(tr_2006_2010, by = c("id" = "id3.2006")) %>% 
    inner_join(eu_members, by = c("id0.2010" = "id0")) %>% 
    rename(id2 = id2.2010, id3 = id3.2010) %>% 
    select(id2, id3, typ_border_A) %>% 
    full_join(tr_2010_2013, by = c("id3" = "id3.2010")) %>% 
    select(id2.2013, id3.2013, typ_border_A) %>% 
    rename(id2 = id2.2013, id3 = id3.2013)
```

```{r border_aggregate, include=FALSE}
# compute percentage of typology per category
border_lvl2 <- typo_border_regions %>% 
    select(id2, typ_border_A) %>% 
    group_by(id2) %>%
    count(typ_border_A)
non_programme <- border_lvl2 %>% group_by(id2) %>% 
    mutate(nr_lvl3 = sum(n)) %>% 
    filter(typ_border_A == 0) %>% 
    mutate(non_programme = round(n/nr_lvl3, 2))
programme <- border_lvl2 %>% group_by(id2) %>% 
    mutate(nr_lvl3 = sum(n)) %>% 
    filter(typ_border_A == 1) %>% 
    mutate(programme = round(n/nr_lvl3, 2))
partially_programme <- border_lvl2 %>% group_by(id2) %>% 
    mutate(nr_lvl3 = sum(n)) %>% 
    filter(typ_border_A == 9) %>% 
    mutate(partially_programme = round(n/nr_lvl3, 2))

# put them together
typ_border_lvl2_2013 <- nuts_2013_levels %>%
    select(id2) %>% 
    full_join(.,non_programme, by = "id2") %>% 
    full_join(., programme, by = "id2") %>%
    full_join(., partially_programme, by = "id2") %>% 
    select(id2, non_programme, programme, partially_programme)
typ_border_lvl2_2013[is.na(typ_border_lvl2_2013)] <- 0.0
```

```{r border_clean, include=FALSE}
# Count records
x_total_cnt <- typ_border_lvl2_2013 %>% 
    count()
# Count valid records
x_2013_cnt <- typ_border_lvl2_2013 %>% 
    filter(non_programme == 0.0 &
               programme == 0.0 &
           partially_programme == 0.0)
# Remove the records that have 0 for each category, because they correspond to the NUTS2006 codes
typ_border_lvl2_clean <- typ_border_lvl2_2013 %>% 
    filter(!(non_programme == 0.0 &
               programme == 0.0 &
           partially_programme == 0.0))
```

```{r border_recode, include=FALSE}
typ_border_clean <- typ_border_lvl2_clean %>% 
    mutate(eligibility = case_when(.$non_programme == 1.0 ~ 0,
                                  .$programme == 1.0 ~ 1,
                                  .$non_programme < 1.0 &
                                      .$programme < 1.0 ~ 2)) %>% 
    rename(id2.2013 = id2, typ_border = eligibility) %>% 
    select(id2.2013, typ_border) %>% 
    mutate(typ_border.incl = 1) %>% 
    distinct()

rm(x_total_cnt, x_2013_cnt, typ_border_lvl2_clean, typo_border_regions, typ_border_lvl2_2013)
```

A sample of the output:

```{r, results="asis", echo=FALSE}
knitr::kable(head(typ_border_clean), title = "A sample of the translated border regions variable")
```

## Variable encoding

**typ_border**

+ **0** – *not-eligible* if the NUTS2 region contains only NUTS 3 regions that are *no programme area*

+ **1** – *eligible* if the NUTS2 region contains only NUTS 3 regions that are *programme area*

+ **2** – *partially* if the NUTS2 region contains NUTS 3 regions from both *no programme area* and *programme area*, or contains only NUTS 3 regions that are *programme area (partly)*

## Data quality

The following aspects might influence the quality of the data:

+ NUTS 3 2006 regions were translated to NUTS 2 2013 regions

```{r clean_up2, include=FALSE}
rm(typo_border_regions, border_lvl2, programme, non_programme,
   partially_programme, x_total_cnt, x_2013_cnt, typ_border_lvl2_clean,
   typ_border_lvl2_2013)
gc()
```
