# Structural Funds eligibility 2007-2013, 2014-2020

**Data source:** Structural Funds eligibility 2007-2013, Structural Funds eligibility 2014-2020 from *European Commission, Directorate-General for Regional and Urban Policy*

**Date of survey:** Not applicable

**Nomenclature:** NUTS 2006, level 2; NUTS 2013, level 2

**Variables used:**

+ Cohesion policy Objective for the NUTS 2 regions, for the programming period 2007-2013

Objective:

    + 1- Convergence
    + 2- Competitiveness
    + 3- Cooperation
    + 4- Multi Objective

+ Eligibility of regions for cohesion funds based on gross domestic product (GDP) per inhabitant (in PPS), by NUTS 2 regions, for the programming period 2014–2020

Classes:

    + 1 – Less developed regions (GDP per inhabitant, < 75)
    + 2 – Transition regions (GDP per inhabitant, >= 75 – < 90)
    + 3 – More developed regions (GDP per inhabitant, >= 90)
    + : – Data not available

## Data processing

1. Convert the relevant variables (objective, classes) to binary variables.

2. Translate the 2007-2013 data to NUTS 2010 then to 2013.

```{r read_struct, include=FALSE, cache=TRUE}
struct_raw_2007 <- read.xlsx(
    "~/Data/COHESIFY/Datasets for regression/SF eligibility/Database of the cumulative allocations to selected projects and expenditure at NUTS2 2007-2013.xlsx",
    sheetIndex = 1, colIndex = c(4,7), startRow = 1, endRow = 17208, 
    header = TRUE, stringsAsFactors = FALSE) %>% 
    rename(nuts_2006 = Location.Cd, class = Objective)
struct_raw_2014 <- read.xlsx(
    "~/Data/COHESIFY/Datasets for regression/SF eligibility/Regional_policies_and_Europe_2020_RYB2016.xlsx",
    sheetIndex = 1, colIndex = c(1,4), startRow = 1, endRow = 287, 
    header = TRUE, stringsAsFactors = FALSE,
    colClasses = c("character", "numeric")) %>% 
    rename(nuts_2013 = NUTS, class = Class)
```

```{r struct_prep_2007, include=FALSE}
# Structural Funds Eligibility 2007-2013, Reshape into binary variables
struct_2007_prep <- struct_raw_2007 %>% 
    select(nuts_2006, class) %>% 
    distinct() %>% 
    arrange(nuts_2006) %>% 
    mutate(class = substr(class, 4, nchar(class))) %>% 
    spread(class, class)

names(struct_2007_prep) <- c("nuts_2006", "Competitiveness", "Convergence",
                             "Cooperation", "Multi.Objective")

struct_2007_prep[!is.na(struct_2007_prep$Competitiveness),
                 "Competitiveness"] <- 1
struct_2007_prep[!is.na(struct_2007_prep$Cooperation),
                 "Cooperation"] <- 1
struct_2007_prep[!is.na(struct_2007_prep$Convergence),
                 "Convergence"] <- 1
struct_2007_prep[!is.na(struct_2007_prep["Multi.Objective"]),
                 "Multi.Objective"] <- 1
struct_2007_prep[is.na(struct_2007_prep)] <- 0

# Translate to 2013
struct_2007_clean <- struct_2007_prep %>% 
    inner_join(tr_2006_2010, by = c("nuts_2006" = "id2.2006")) %>% 
    select(id2.2010, Competitiveness, Convergence, Cooperation, Multi.Objective) %>% 
    distinct() %>% 
    inner_join(tr_2010_2013, by = "id2.2010") %>% 
    select(id2.2013, Competitiveness, Convergence, Cooperation, Multi.Objective) %>% 
    distinct() %>% 
    mutate(struct_2007.incl = 1)

```

```{r struct_2007_test, include=FALSE, eval=FALSE}
# Testing: cooperation and multi-objective are not mutually exclusive, 
# convergence and competitiveness are mutually exclusive
struct_2007_prep %>% 
    filter(Convergence == 1 & Competitiveness == 1) %>% 
    arrange(nuts_2006)

# struct_raw_2007 %>% 
#     filter(nuts_2006 == "HU10" | nuts_2006 == "CZ01") %>% 
#     distinct() %>% 
#     arrange(nuts_2006)
```

```{r struct_prep_2014, include=FALSE}
# Structural Funds Eligibility 2014-2020, Reshape into binary variables
struct_2014_prep <- struct_raw_2014 %>% 
    select(nuts_2013, class) %>% 
    distinct() %>% 
    arrange(nuts_2013) %>% 
    mutate(class = case_when(.$class == 1 ~ "Less.Developed",
                             .$class == 2 ~ "Transition",
                             .$class == 3 ~ "More.Developed")) %>% 
    spread(class, class) %>% 
    select(nuts_2013, Less.Developed, Transition, More.Developed) %>% 
    inner_join(nuts_2013_levels, by = c("nuts_2013" = "id2")) %>% 
    select(-id0, -id1, -id3) %>% 
    rename(id2.2013 = nuts_2013) %>% 
    distinct()

struct_2014_prep[!is.na(struct_2014_prep$Less.Developed),
                 "Less.Developed"] <- 1
struct_2014_prep[!is.na(struct_2014_prep$Transition),
                 "Transition"] <- 1
struct_2014_prep[!is.na(struct_2014_prep$More.Developed),
                 "More.Developed"] <- 1
struct_2014_prep[is.na(struct_2014_prep)] <- 0

# Testing: all 3 categories are mutually exclusive
struct_2014_prep %>% 
    filter(Less.Developed == 1 & Transition == 1)
struct_2014_prep %>% 
    filter(Less.Developed == 1 & More.Developed == 1)
struct_2014_prep %>% 
    filter(More.Developed == 1 & Transition == 1)

struct_2014_clean <- struct_2014_prep %>% 
    mutate(struct_2014.incl = 1)
rm(struct_2014_prep)
```

A sample of the output:

```{r, results='asis', echo=FALSE}
knitr::kable(head(struct_2007_clean), caption = "Structural Funds eligibility 2007-2013")
```

```{r, results='asis', echo=FALSE}
knitr::kable(head(struct_2014_clean), caption = "Structural Funds eligibility 2014-2020")
```

## Variable encoding

For programming period 2007-2013:

**Competitiveness:** `1` – region is part of the objective *Competitiveness*, or not `0`

**Convergence:** `1` – region is part of the objective *Convergence*, or not `0`

**Cooperation:** `1` – region is part of the objective *Cooperation*, or not `0`

**Multi.Objective:** `1` – region is part of the objective *Multi Objective*, or not `0`

For porgramming period 2014-2020:

**Less.Developed:** `1` – region is part of the class *Less developed regions (GDP per inhabitant, < 75)*, or not `0`

**Transition:** `1` – region is part of the class *Transition regions (GDP per inhabitant, >= 75 – < 90)*, or not `0`

**More.Developed:** `1` – region is part of the class *More developed regions (GDP per inhabitant, >= 90)*, or not `0`


## Data quality

The following aspects might influence the quality of the data:

+ NUTS 2 2006 regions were translated to NUTS 2 2013 regions

+ The NUTS 2013 region *FI1D* is a merge of the NUTS 2006 regions *FI13* and *FI1A*. Both of these 2006 regions were part of the *Competitiveness* objective, but *FI1A* was also part of the *Cooperation* objective. Therefore after version translation the region *FI1D* has both `0` and `1` for the *Cooperation* variable. Similarly, the NUTS 2013 region *DED4* is a merge of the NUTS 2006 regions *DED1* and *DED3* where *DED1* was part of the *Cooperation* objective while *DED3* was not. In both cases the corresponding NUTS 2013 region recieves `1` for the *Cooperation* variable and the record of `0` *Cooperation* is deleted.

```{r clean_up5, include=FALSE, message=FALSE, warning=FALSE}
rm(struct_2007_prep, struct_raw_2007, struct_raw_2014)
gc()
```
