# Unemployment

**Data source:** ESPON Database Portal – Europe 2020 Index based on the distances to 4 of the 8 Europe 2020 Strategy

**Date of survey:** 2013

**Date of access:** 2016-12-02 12:32:59

**Nomenclature:** NUTS 2013 level 2

**Variables used:**

+ Unemployment rate of persons aged 15-74 (`lfst_r_lfu3rt`)

## Data processing

1. Select values for NUTS2 regions

```{r unemp_load, include=FALSE, cache=TRUE}
unemp_clean <- read.xlsx(
    "~/Data/COHESIFY/Datasets for regression/Employment and poverty/Unemployment rates by sex, age and NUTS 2 regions (_) lfst_r_lfu3rt.xls",
    sheetIndex = 1, colIndex = c(1,4), startRow = 11, endRow = 502, 
    header = TRUE, stringsAsFactors = FALSE,
    colClasses = c("character", "numeric")) %>% 
    rename(id2.2013 = GEO.TIME, unemployment_2013 = X2013) %>% 
    filter(nchar(id2.2013) >= 4) %>% 
    distinct() %>% 
    inner_join(nuts_2013_levels, by = c("id2.2013" = "id2")) %>% 
    inner_join(eu_members, by = "id0") %>% 
    select(id2.2013, unemployment_2013) %>% 
    distinct() %>% 
    mutate(unemp.incl = 1)
```

A sample of the output:

```{r, results='asis', echo=FALSE}
knitr::kable(head(unemp_clean), caption = "Unemployment rate of persons aged 15-74, NUTS 2013 level 2")
```

## Variable encoding

**unemployment_2013:** Unemployment rate of persons aged 15-74

## Data quality

No potential errors were introduced during the processing.
