# Social progress (Basic Human Needs, Wellbeing, Opportunity)

**Data source:** EU regional Social Progress Index (SPI) 2016 [http://www.socialprogressimperative.org/custom-indexes/european-union/](http://www.socialprogressimperative.org/custom-indexes/european-union/)

**Date of survey/preparation:** 2016

**Nomenclature:** NUTS 2010 level 2

**Variables used:**

+ European Union Regional Social Progress Index

## Data processing

1. Translate from NUTS 2 2010 to NUTS 2 2013

```{r soc_load, include=FALSE, cache=TRUE}
soc_clean <- read.xlsx(
    "~/Data/COHESIFY/Datasets for regression/Social Progress Index/2016_eu_spi.xlsx",
    sheetIndex = 1, colIndex = c(1,4), startRow = 1, endRow = 273, 
    header = TRUE, stringsAsFactors = FALSE,
    colClasses = c("character", "numeric")) %>% 
    rename(nuts_2010 = NUTS.Code.2010,
           soc_idx = European.Union.Regional.Social.Progress.Index) %>% 
    inner_join(tr_2010_2013, by = c("nuts_2010" = "id2.2010")) %>% 
    inner_join(eu_members, by = c("id0.2013" = "id0")) %>% 
    select(id2.2013, soc_idx) %>% 
    distinct() %>% 
    mutate(soc_idx.incl = 1)
```

A sample of the output:

```{r, results='asis', echo=FALSE}
knitr::kable(head(soc_clean), caption = "Social progress, NUTS 2013 level 2")
```

## Variable encoding

**soc_idx:** European Union Regional Social Progress Index

## Data quality

The following aspects might influence the quality of the data:

+ NUTS 2 2010 regions were translated to NUTS 2 2013 regions