# Access to social services

**Data source:** Eurostat / European Atlas of Services of General Interest / Social Services of General Interest

**Date of survey:** NA

**Nomenclature:** NUTS 2006 level 2

**Variables used:**

+ SGI-aggregated

## Data processing

1. The record for *Saarland* was missing its NUTS code thus it was manually added.

2. Translate NUTS 2006 to 2010 then to 2013.

```{r acc2soc_load, include=FALSE, cache=TRUE}
acc2soc_raw <- read.xlsx(
    "~/Data/COHESIFY/Datasets for regression/Access to services/SGI typology.xlsx",
    sheetIndex = 1, colIndex = c(1,7), startRow = 1, endRow = 287, 
    header = TRUE, stringsAsFactors = FALSE) %>% 
    rename(nuts_2006 = NUTS2, sgi_aggregated = SGI.aggreg.)
```

```{r acc2soc_tranlate, include=FALSE}
acc2soc_translate <- acc2soc_raw %>% 
    inner_join(tr_2006_2010, by = c("nuts_2006" = "id2.2006")) %>% 
    rename(id2.2006 = nuts_2006) %>% 
    select(id2.2010, sgi_aggregated) %>% 
    distinct() %>% 
    mutate(x = substr(id2.2010, 0, 2)) %>% 
    inner_join(eu_members, by = c("x" = "id0"))

acc2soc_clean <- acc2soc_translate %>% 
    inner_join(tr_2010_2013, by = "id2.2010") %>% 
    select(id2.2013, sgi_aggregated) %>% 
    distinct() %>% 
    mutate(sgi_aggregated.incl = 1)
```

```{r check_acc2soc, include=FALSE}
# Check if all EU members are present, should give 28
acc2soc_translate %>%
    mutate(x = substr(id2.2010, 0, 2)) %>% 
    select(x) %>% 
    distinct()
```

A sample of the output:

```{r, results='asis', echo=FALSE}
knitr::kable(head(acc2soc_clean), caption = "Access to social services, NUTS 2013 level 2")
```

## Variable encoding

**sgi_aggregated:** aggregated social services index

## Data quality

The following aspects might influence the quality of the data:

+ NUTS 2 2006 regions were translated to NUTS 2 2013 regions.

+ The NUTS0 region *HR* is not present in the data.

```{r clean_up7, include=FALSE, message=FALSE, warning=FALSE}
rm(acc2soc_translate, acc2soc_raw)
gc()
```

