# Access to social services

**Data source:** Eurostat / European Atlas of Services of General Interest / Social Services of General Interest

**Date of survey:** NA

**Nomenclature:** NUTS 2006 level 2

**Variables used:**

+ SGI-aggregated

## Data processing

1. The record for *Saarland* was missing its NUTS code thus it was manually added.

2. Translate NUTS 2006 to 2010 then to 2013.

```{r acc2soc_load, include=FALSE, cache=TRUE}
acc2soc_raw <- read.xlsx(
    "~/Data/COHESIFY/Datasets for regression/Access to services/SGI typology_2.xlsx",
    sheetIndex = 1, colIndex = c(1,3,6), startRow = 1, endRow = 287, 
    header = TRUE, stringsAsFactors = FALSE) %>% 
    rename(nuts_2006 = NUTS2, sgei = SGEI.aggreg., ssgi = SSGI.aggreg.)
```

```{r acc2soc_tranlate, include=FALSE}
acc2soc_clean <- acc2soc_raw %>% 
    inner_join(distinct(select(tr_2006_2010, id2.2010, id2.2006)),
               by = c("nuts_2006" = "id2.2006")) %>% 
    rename(id2.2006 = nuts_2006) %>% 
    inner_join(
        distinct(select(tr_2010_2013, -id3.2010, -id3.2013)),
        by = "id2.2010"
    ) %>% 
    select(id2.2013, sgei, ssgi)
```

The following regions were merged, split or whatever in the previous NUTS versions, thus it is not possible now to clearly determine a value for them. Therefore there are set a missing value.

```{r}
acc2soc_clean[duplicated(acc2soc_clean$id2.2013), "id2.2013"]
```

```{r}
# see comment at Data quality
acc2soc_clean[acc2soc_clean$id2.2013 == "FI1D", c("sgei", "ssgi")] <- NA
acc2soc_clean[acc2soc_clean$id2.2013 == "DE40", c("sgei", "ssgi")] <- NA
acc2soc_clean[acc2soc_clean$id2.2013 == "DED4", c("sgei", "ssgi")] <- NA
acc2soc_clean[acc2soc_clean$id2.2013 == "SI03", c("sgei", "ssgi")] <- NA

acc2soc_clean <- filter(acc2soc_clean, !is.na(sgei))
```


```{r check_acc2soc, include=FALSE}
# Check if all EU members are present, should give 28
acc2soc_translate %>%
    mutate(x = substr(id2.2010, 0, 2)) %>% 
    select(x) %>% 
    distinct()
```

A sample of the output:

```{r, results='asis', echo=FALSE}
knitr::kable(head(acc2soc_clean), caption = "Access to social services, NUTS 2013 level 2")
```

## Variable encoding

**sgi_aggregated:** aggregated social services index

## Data quality

The following aspects might influence the quality of the data:

+ NUTS 2 2006 regions were translated to NUTS 2 2013 regions.

+ The NUTS0 region *HR* is not present in the data.

+ Merge of *FI1A* + *FI13* = *FI1D*. Merge of *DE41* + *DE42* = *DE40*. In both cases the NUTS 2013 regions (*FI1D*, *DE40*) are assigned a missing value, because population data is not available for NUTS 2006 regions, thus a weighted average value for the NUTS 2010/2013 region cannot be computed. Also *DED4* and *SI03*.

```{r clean_up7, include=FALSE, message=FALSE, warning=FALSE}
rm(acc2soc_translate, acc2soc_raw)
gc()
```

