# Share of population employed in agriculture

**Data source:** Eurostat

**Last update:** 24.01.2017

**Date of access:** 21.02.2017

**Nomenclature:** NUTS 2013

**Variable used:**

+ Employment in technology and knowledge-intensive sectors by NUTS 2 regions and sex (from 2008 onward, NACE Rev. 2) [`htec_emp_reg2`] / Agriculture, forestry and fishing; mining and quarrying [Percentage of total employment]

## Data processing

1. Parse NUTS2 region codes from code-name combinations.

2. Input missing values for NUTS2 regions from higher level regions. Depending on the availability, either from NUTS1 or NUTS0. 

```{r agr_load, include=FALSE, cache=TRUE}
agr_raw <- read.xlsx(
    "~/Data/COHESIFY/Datasets for regression/Agriculture/htec_emp_reg2.xls",
    sheetIndex = 1, colIndex = c(1,7), startRow = 1442, endRow = 1910, 
    header = TRUE, stringsAsFactors = FALSE,
    colClasses = c("character", "numeric")) %>% 
    rename(x = GEO.TIME,
           agr_share = X2013)
```

```{r agr_prep, include=FALSE}
# parse the nuts id from the names
agr_clean <- agr_raw
agr_clean$nuts_2013 <- unlist(lapply(strsplit(agr_raw$x, " "), function(x) x[1]))
agr_clean <- agr_clean %>% select(-x)

agr_lvl0 <- agr_clean %>% filter(nchar(nuts_2013) == 2) %>% 
    rename(id0 = nuts_2013)
agr_lvl1 <- agr_clean %>% filter(nchar(nuts_2013) == 3) %>% 
    mutate(id0 = substr(nuts_2013, 0, 2)) %>% 
    rename(id1 = nuts_2013)
agr_lvl2 <- agr_clean %>% filter(nchar(nuts_2013) == 4) %>% 
    mutate(id0 = substr(nuts_2013, 0, 2)) %>% 
    mutate(id1 = substr(nuts_2013, 0, 3)) %>% 
    rename(id2 = nuts_2013)

# put all the values for all regions together;
# regions are only included in the strict analyisis if they have a 
# value for the nuts2 region
agr_all <- agr_lvl0 %>% 
    full_join(agr_lvl1, by = "id0", suffix = c(".id0", ".id1")) %>% 
    full_join(agr_lvl2, by = "id1") %>% 
    rename(agr_share.id2 = agr_share, id0 = id0.x) %>% 
    select(-id0.y) %>%
    inner_join(eu_members, by = "id0") %>% 
    inner_join(select(nuts_2013_levels, id2), by = "id2") %>% 
    distinct() %>% 
    # mutate(agr.incl = case_when(!is.na(.$agr_share.id2) ~ 1,
    #                             TRUE ~ 0)) %>%
    full_join(nuts_2013_single_n0, by = "id0") %>% 
    full_join(nuts_2013_single_n1, by = "id1") %>% 
    mutate(agr.incl = case_when(.$incl.x == 1 |
                                (.$incl.y == 1 & !is.na(.$agr_share.id1)) |
                                !is.na(.$agr_share.id2) ~ 1)) %>% 
    arrange(id0.x)

# Input country average to missing NUTS2 values
missing <- agr_all %>% filter(is.na(agr.incl))

for (r in missing$id2) {
    # if NUTS1 value missing
    if (is.na(agr_all[agr_all$id2 == r, "agr_share.id1"])) {
        agr_all[agr_all$id2 == r,
                "agr_share.id2"] <- agr_all[agr_all$id2 == r, "agr_share.id0"]
    } else {
    # NUTS1 value present
        agr_all[agr_all$id2 == r,
                "agr_share.id2"] <- agr_all[agr_all$id2 == r, "agr_share.id1"]
    }
}

agr_clean <- agr_all %>%
    select(id2, agr_share.id2, agr.incl) %>% 
    rename(agr_share = agr_share.id2, id2.2013 = id2)

agr_clean[is.na(agr_clean)] <- 0
```

A sample of the output:

```{r, results='asis', echo=FALSE}
knitr::kable(head(agr_clean), caption = "Share of population employed in agriculture, NUTS 2013 level 2")
```

## Variable encoding

**agr_share:** Agriculture, forestry and fishing; mining and quarrying [Percentage of total employment]

## Data quality

The following aspects might influence the quality of the data:

+ The NUTS2 regions below are missing values. For these regions the corresponding NUTS1 value is imputed if available, otherwise the corresponding NUTS0 value.

The NUTS2 regions that needed value imputation from higher level regions are:


```{r agr_missing_table, results='asis', echo=FALSE}
unique(agr_clean[agr_clean$agr.incl == 0, "id2.2013"])
# knitr::kable(select(missing, agr_share.id0, id0.x, agr_share.id1, id1), caption = "NUTS2 regions with missing agriculture index values")
```

```{r clean_up71, include=FALSE, message=FALSE, warning=FALSE}
rm(agr_raw, agr_all, agr_lvl2, agr_lvl1, agr_lvl0, missing, r)
gc()
```

