# EU Identification

**Data source:** Internal

**Date of survey:** NA

**Nomenclature:** NUTS 2010 level 2

**Variables used:**

+ EU identification

    + 11 – Negative - Neutral
    + 12 – Negative - Attached
    + 21 – Neutral - Attached
    + 22 – Neutral - Neutral
    + 31 – Positive

## Data processing

1. Translate NUTS 2 2010 to 2013 

```{r eu_id_load, include=FALSE}
eu_id_clean <- read.xlsx(
    "~/Data/COHESIFY/Datasets for regression/EU Identification/EU Identification typology (COHESIFY Output 2.4a).xlsx",
    sheetIndex = 1, colIndex = c(1,3,4), startRow = 1, endRow = 273, 
    header = TRUE, stringsAsFactors = FALSE,
    colClasses = c("character", "numeric", "character")) %>% 
    rename(id2.2010 = NUTS_2_.2010., eu_id_type_code = Type_Code,
           eu_id_type = EU_Id_Type) %>% 
    inner_join(
        filter(
            select(tr_2010_2013, -id3.2010, -id3.2013),
            !(id2.2010 == "SI02" & id2.2013 == "SI03")
            ),
        by = "id2.2010") %>% 
    select(id2.2013, eu_id_type_code) %>% 
    distinct() %>% 
    mutate(eu_id.incl = case_when(!is.na(.$eu_id_type_code) ~ 1,
                                  TRUE ~ 0))

missing <- eu_id_clean %>% 
    filter(is.na(eu_id_type_code)) %>% 
    select(id2.2013) %>% 
    distinct()
```

## Variabale encoding

**eu_id_type_code:** EU identification as given above

## Data quality

The following aspects might influence the quality of the data:

+ NUTS2 regions with missing values

```{r eu_id_missing, results='asis', echo=FALSE}
knitr::kable(missing, caption = "NUTS2 regions with missing values for the variable EU identification")
```

```{r clean_up8, include=FALSE, message=FALSE, warning=FALSE}
rm(missing)
```

