# EU Identification

**Data source:** Internal

**Date of survey:** NA

**Nomenclature:** NUTS 2010 level 2

**Variables used:**

+ EU identification

    + Positive [%] – percentage of respondents having a positive image about the EU
    + Neutral [%] – percentage of respondents having a neutral image about the EU
    + Negative [%] – percentage of respondents having a negative image about the EU
    + DK [%] – percentage of respondents who has no opinion

## Data processing

1. Translate NUTS 2 2010 to 2013 

2. Missing values in the input data are represented as empty strings or with the value `-1`. These are recoded as `NA`.

```{r eu_id_load, include=FALSE}
eu_id <- read.xlsx(
    "~/Data/COHESIFY/Datasets for regression/EU Identification/EU Image Percentages.xlsx",
    sheetIndex = 1, colIndex = c(1,2,3,4,5), startRow = 1, endRow = 286, 
    header = TRUE, stringsAsFactors = FALSE,
    colClasses = c("character", "numeric", "numeric","numeric","numeric")) %>% 
    rename(id2.2010 = NUTS_Code, eu_id_positive = Positive,
           eu_id_neutral = Neutral, eu_id_negative = Negative,
           eu_id_dk = DK) %>% 
    inner_join(
        filter(
            select(tr_2010_2013, -id3.2010, -id3.2013),
            !(id2.2010 == "SI02" & id2.2013 == "SI03")
            ),
        by = "id2.2010") %>% 
    select(id2.2013, eu_id_positive, eu_id_neutral, eu_id_negative,
           eu_id_dk) %>% 
    distinct()
# replace -1 with NA
eu_id[eu_id==-1] <- NA
# mark regions with missing values
eu_id_clean <- eu_id %>% 
    mutate(eu_id.incl = case_when(complete.cases(.) ~ 1,
                                  TRUE ~ 0))

missing <- eu_id_clean %>% 
    filter(eu_id.incl == 0) %>% 
    select(id2.2013) %>% 
    distinct()
```

## Variabale encoding

**eu_id_positive:** percentage of respondents having a positive image about the EU

**eu_id_neutral:** percentage of respondents having a neutral image about the EU

**eu_id_positive:** percentage of respondents having a negative image about the EU

**eu_id_positive:** percentage of respondents who has no opinion


## Data quality

The following aspects might influence the quality of the data:

+ NUTS2 regions with missing values

```{r eu_id_missing, results='asis', echo=FALSE}
knitr::kable(missing, caption = "NUTS2 regions with missing values for the variable EU identification")
```

```{r clean_up8, include=FALSE, message=FALSE, warning=FALSE}
rm(missing, eu_id)
```

