# EU Attachment

**Data source:** Internal

**Date of survey:** NA

**Nomenclature:** NUTS 2010 level 2

**Variables used:**

+ EU identification

    + Attached_P [%] – percentage of respondents feeling attached to the EU
    + Not_Very_P [%] – percentage of respondents feeling not very attached to the EU
    + Not_At_All_P [%] – percentage of respondents feeling not attached at all to the EU
    + DK_P [%] – percentage of respondents who has no opinion

# Data processing

1. Translate NUTS 2 2010 to 2013 

2. Missing values in the input data are represented as `NULL`. There are also `0` values, but these are not treated as missing value.

```{r eu_attach_load, include=FALSE}
eu_att_clean <- read.xlsx(
    "~/Data/COHESIFY/Datasets for regression/EU Attachment/FINAL_Merged_84.3_84.1_ATTACHMENT_EU_ALL_NUTS_LEVELS_Percentages.xlsx",
    sheetIndex = 2, colIndex = c(1,2,3,4,5), startRow = 1, endRow = 273, 
    header = TRUE, stringsAsFactors = FALSE,
    colClasses = c("character", "numeric", "numeric","numeric","numeric")) %>% 
    rename(id2.2010 = NUTS_2, eu_att_positive = Attached_P,
           eu_att_neutral = Not_Very_P, eu_att_negative = Not_At_All_P,
           eu_att_dk = DK_P) %>% 
    inner_join(
        filter(
            select(tr_2010_2013, -id3.2010, -id3.2013),
            !(id2.2010 == "SI02" & id2.2013 == "SI03")
            ),
        by = "id2.2010") %>% 
    select(id2.2013, eu_att_positive, eu_att_neutral, eu_att_negative,
           eu_att_dk) %>% 
    distinct() %>% 
    mutate(eu_att.incl = case_when(complete.cases(.) ~ 1,
                                  TRUE ~ 0))
# select regions with missing values
missing <- eu_att_clean %>% 
    filter(eu_att.incl == 0) %>% 
    select(id2.2013) %>% 
    distinct()
```

## Variabale encoding

**eu_att_positive:** percentage of respondents feeling attached to the EU

**eu_att_neutral:** percentage of respondents feeling not very attached to the EU

**eu_att_negative:** percentage of respondents feeling not attached at all to the EU

**eu_att_dk:** percentage of respondents who has no opinion

## Data quality

The following aspects might influence the quality of the data:

+ NUTS2 regions with missing values

```{r eu_att_missing, results='asis', echo=FALSE}
knitr::kable(missing, caption = "NUTS2 regions with missing values for the variable EU attachment")
```

```{r clean_up9, include=FALSE, message=FALSE, warning=FALSE}
rm(missing)
```